---
output: html_document
editor_options: 
  chunk_output_type: console
---
class: inverse

# Poll

Have you ever heard of tidy evaluation?

* No

* I've heard of it, but I'm not sure what it is

* I know what it is, but I'm not comfortable using it

* Yes, I feel comfortable using it

---

# Goals for Today

Be able to write a function that utilizes `{tidyverse}` functions (e.g., 
`mutate`, `select`, `filter`, etc.) by using tidy evaluation

---

# Outline 

* Function Basics

* Nuances of writing functions with the tidyverse

  - data-masking
  
  - tidy-select

* Writing functions that use data-masking tidyverse functions

* Writing functions that use tidy-select tidyverse functions

* Practice

* Resources & Additional Material

---

# Functions in R

**Rule of thumb**: If you're copy/pasting more than twice, write a function!

Functions:

* Prevent inconsistencies because they force multiple computations to follow a
single recipe

* Emphasize what varies (arguments) and what is constant (everything else)

* Make updates easier since you only modify one place

* Make code clearer, especially if you give function and arguments informative names

If you're a `{tidyverse}` user, you'll likely want to use the `{tidyverse}` for
your functions.

However, because of some technical features of the `{tidyverse}` (*data masking*),
there are a few nuances to writing tidyverse-style functions. 

---

# Anatomy of a Function

```{r, eval=FALSE}
function_name <- function(argument1, argument2){
  # body of the function
  # code to repeat
}

# the function call
function_name(argument1 = "arg1",
              argument2 = "arg2")
```

---

# What's Returned by a Function

* By default, only the last known operation is returned by the function

* This is *different* from SAS macros

* Can use a named list to return multiple objects

```{r, eval=FALSE}
example_return_list <- function(arg1){
  # function body
  
  # create object 1
  
  # create object 2
  
  return(list("object1" = object1,
              "object2" = object2))
}

# store the function call 
call_return_list <- example_return_list(arg1)

# will return
call_return_list$object1
call_return_list$object2
```

---

# Steps for Writing a Function

1. Write and test your code in R, outside of a function

2. Enclose in `function(){[working code here]}` with an informatively named 
function and arguments

3. Adapt code to replace variable names and/or values with function arguments,
as needed

