class: inverse, center, middle
# Writing tidyverse-style Functions

---
# What is data-masking?

.pull-left[
* Data-masking is a distinctive feature of R whereby programming is performed directly on a data set, with columns defined as normal objects.
]

.pull-right[
* While data-masking makes it easy to program interactively with data frames, it makes it harder to create functions.
]

```{r, eval = FALSE}
# Unmasked programming
mean(mtcars$cyl + mtcars$am)
#> [1] 6.59375

# Referring to columns is an error - Where is the data?
mean(cyl + am)
#> Error in mean(cyl + am): object 'cyl' not found

# Data-masking
with(mtcars, mean(cyl + am))
#> [1] 6.59375
```

---
# Why do I need `{{ · }}`?

* Passing data-masked arguments to functions requires injection (also known as quasiquotation) with the embracing operator `{{ · }}` or, in more complex cases, the injection operator `!!`.

* This is needed because under the hood data-masking works by defusing R code to prevent its immediate evaluation.

* The defused code is resumed later on in a context where data frame columns are defined.

```{r, eval = FALSE}
my_mean <- function(data, var1, var2) {
  dplyr::summarise(data, mean(var1 + var2))
}

my_mean(mtcars, cyl, am)
#> Error in `dplyr::summarise()`:
#> ! Problem while computing `..1 = mean(var1 + var2)`.
#> Caused by error in `mean()`:
#> ! object 'cyl' not found
```

???

- The problem here is that `summarise()` defuses the R code it was supplied, i.e. mean(var1 + var2). 

- Instead we want it to see mean(cyl + am). 

- This is why we need injection, we need to modify that piece of code by injecting the code supplied to the function in place of var1 and var2.

---
# How to use `{{ · }}`?

* To inject a function argument in data-masked context, just embrace it with `{{ · }}`

```{r, eval = FALSE}
my_mean <- function(data, var1, var2) {
  dplyr::summarise(data, mean({{ var1 }} + {{ var2 }}))
}

my_mean(mtcars, cyl, am)
#> # A tibble: 1 x 1
#>   `mean(cyl + am)`
#>              <dbl>
#> 1             6.59
```

---
# Alternatives to `{{ · }}`?

* `{{ · }}` is a shortcut for `!!enquo(·)`

* Use the `!!` and `rlang::enquo(·)` combination when you need to pass `var1` and `var2` around before it's injected.

```{r, eval = FALSE}
my_mean <- function(data, var1, var2) {
  var1_quo <- rlang::enquo(var1)
  var2_quo <- rlang::enquo(var2)
  
  dplyr::summarise(data, mean(!!var1_quo + !!var2_quo))
}

my_mean(mtcars, cyl, am)
#> # A tibble: 1 x 1
#>   `mean(cyl + am)`
#>              <dbl>
#> 1             6.59
```

---
# Can I avoid `{{ · }}` and `!!enquo(·)`?

* YES! (mostly)

* Immediately convert inputs into column name **strings**, and utilize all the what you learned in the first portion of this presentation.

```{r, eval = FALSE}
my_mean <- function(data, var1, var2) {
  var1 <- dplyr::select(data, {{ var1 }}) |> names()
  var2 <- dplyr::select(data, {{ var2 }}) |> names()
  
  dplyr::summarise(data, mean(.data[[var1]] + .data[[var2]]))
}

my_mean(mtcars, cyl, am)
#> # A tibble: 1 x 1
#>   `mean(.data[["cyl"]] + .data[["am"]])`
#>                                    <dbl>
#> 1                                   6.59
```