---
output: html_document
editor_options: 
  chunk_output_type: console
---
class: inverse, center, middle
# Let's Practice

---

# Practice with {ggplot}

Write a function that uses `{ggplot}` to create a scatterplot of age versus
marker value from the `gtsummary::trial` dataset, where you can vary:

* The subset of the data that is plotted (e.g., trt = Drug A or age > 65)

* The variable used to color code the points (stage or grade)

```{r, eval=FALSE}
head(gtsummary::trial)
# A tibble: 6 x 8
#   trt      age marker stage grade response death ttdeath
#   <chr>  <dbl>  <dbl> <fct> <fct>    <int> <int>   <dbl>
# 1 Drug A    23  0.16  T1    II           0     0    24  
# 2 Drug B     9  1.11  T2    I            1     0    24  
# 3 Drug A    31  0.277 T1    II           0     0    24  
# 4 Drug A    NA  2.07  T3    III          1     1    17.6
# 5 Drug A    51  2.77  T4    III          1     1    16.4
# 6 Drug B    39  0.613 T4    I            0     1    15.6
```

---
# Step 1

Write ggplot code that works and is close to what we want

---

# Step 1

Write ggplot code that works and is close to what we want

```{r, fig.height=3}
library(ggplot2)

gtsummary::trial %>% 
  dplyr::filter(trt == "Drug A") %>% 
  ggplot(aes(x = age,
             y = marker, 
             color = stage)) +
  geom_point()
```

---

# Step 2

Put the working code into a function (haven't yet used function arguments)

---

# Step 2

Put the working code into a function (haven't yet used function arguments)

```{r, eval=FALSE}
plot_age_marker <- function(filter_condition, color_variable){
  gtsummary::trial %>% 
    dplyr::filter(trt == "Drug A") %>% 
    ggplot(aes(x = age,
             y = marker, 
             color = stage)) +
    geom_point()
}
```

---

# Step 3

Replace the value of stage with the color_variable function input

---

# Step 3

Replace the value of stage with the color_variable function input

```{r}
plot_age_marker <- function(filter_condition, color_variable){
  gtsummary::trial %>% 
    dplyr::filter(rlang::eval_bare(rlang::parse_expr(filter_condition))) %>% 
    ggplot(aes(x = age,
             y = marker, 
             color = .data[[color_variable]])) +
    geom_point()
}
```

---

# Try to run our function

```{r, fig.height=4}
# try to run our function
plot_age_marker(filter_condition = "age > 40",
                color_variable = "stage")
```

---

# Try to run our function

Now, what if we want to subset the data based on a categorical variable?

```{r, eval = FALSE}
plot_age_marker(filter_condition = 'trt == "Drug A"',
                color_variable = "stage")
```

```{r, fig.height=3}
plot_age_marker(filter_condition = "trt == \"Drug A\"",
                color_variable = "stage")
```

