---
output: html_document
editor_options: 
  chunk_output_type: console
---

class: inverse, center, middle

# Additional Material

---

# rlang

* Examples for when you can't use the walrus operator and have to use
`rlang::sym()` or `rlang::syms()`

* `rlang::sym()` is used when you have a string and you want it to be unquoted

* `rlang::syms()` is used when you have a vector and you want it to be unquoted

```{r, eval=FALSE}
# e.g., a vector of c("A", "B", "C")

vec <- c("mpg", "hp")

# Error: Can't subset columns that don't exist.
# x Column `vec` doesn't exist.
mtcars %>% 
  dplyr::select(vec)

# now works
# similar to `all_of()` but there are times when you need the below
mtcars %>% 
  dplyr::select(!!!rlang::syms(vec))
```

* Can also be used to name variables within functions

---

# rlang

* Can use `rlang::expr()` to see what `rlang::sym()` is evaluating to

```{r, eval=FALSE}
vec <- c("mpg", "hp")

rlang::expr(mtcars %>% 
  dplyr::select(!!!rlang::syms(vec)))
# mtcars %>% dplyr::select(mpg, hp)
```

---

# rlang: unquoted expresssions

* You can also write functions that accept unquoted expressions as arguments

```r
my_filter <- function(data, condition) {
  condition <- rlang::enquo(condition)
  
  dplyr::filter(data, !!condition)
}

my_filter(gtsummary::trial, age < 30) |> head()
#> # A tibble: 6 x 8
#>   trt      age marker stage grade response death ttdeath
#>   <chr>  <dbl>  <dbl> <fct> <fct>    <int> <int>   <dbl>
#> 1 Drug A    23  0.16  T1    II           0     0    24  
#> 2 Drug B     9  1.11  T2    I            1     0    24  
#> 3 Drug B    21  0.258 T4    I            0     1    12.9
#> 4 Drug B    28  0.803 T4    II           0     1    18  
#> 5 Drug B    25  2.45  T1    I            1     0    24  
#> 6 Drug B    25  0.531 T4    III          0     1    23.2

```