---
output: html_document
editor_options: 
  chunk_output_type: console
---

class: inverse, center, middle

# Additional Material

---

# Writing functions that involve base R and the `{tidyverse}`

---

# rlang

* Examples for when you can't use the walrus operator and have to use
`rlang::sym()` or `rlang::syms()`

* `rlang::sym()` is used when you have a string and you want it to be unquoted

* `rlang::syms()` is used when you have a vector and you want it to be unquoted

```{r, eval=FALSE}
# e.g., a vector of c("A", "B", "C")

vec <- c("mpg", "hp")

# Error: Can't subset columns that don't exist.
# x Column `vec` doesn't exist.
mtcars %>% 
  dplyr::select(vec)

# now works
# similar to `all_of()` but there are times when you need the below
mtcars %>% 
  dplyr::select(!!!rlang::syms(vec))
```

* Can also be used to name variables within functions

* Can use `rlang::expr()` to see what `rlang::sym()` is evaluating to

```{r, eval=FALSE}
vec <- c("mpg", "hp")

rlang::expr(mtcars %>% 
  dplyr::select(!!!rlang::syms(vec)))
# mtcars %>% dplyr::select(mpg, hp)
```

